---
title: "DATAFRAMES IN R - PART 01"
author: "PRAMISH TM"
date: "2024-11-26"
output: html_document
---


# Why use DataFrames?
- Ease of usage
- Functionalities
- Most data comes in tabular format
- compatible with analytical tools

# Creating Data.Frame

```{r}
country <- c("France", "Germany", "Greece", "Italy", "Portugal", "Spain")
```

```{r}
gdp_growth <- c(1.3, 0.3, 1.9, 0.3, NA, 2)
```
```{r}
mkt_type <- factor(c("High", "High", "Low", "Medium", "Low", "Medium"))
```

```{r}
print(country)
```
```{r}
print(gdp_growth)
```
```{r}
print(mkt_type)
```
# Let's convert these vectors into a Data.Frame now

```{r}
data <- data.frame(country = country, gdp_growth = gdp_growth, mkt_type = mkt_type)
print(data)
```
```{r}
class(data)
```

```{r}
View(data)
```

```{r}
edit(data)
```
# Various dataframe operations

## Accessing the individual columns using the "$" notation

```{r}
data
```

```{r}
data$country
```
```{r}
print(data$country)
```

```{r}
print(data$gdp_growth)
```
```{r}
print(data$mkt_type)
```
- It is to be noted that when we type a space in the column names, it is replaced by a "." in the dataframe by default.

# Creating new columns

```{r}
data$abbreviated_country <- abbreviate(data$country, minlength = 3)
```
```{r}
data$abbreviated_country
```
```{r}
data
```

## Structure of a dataframe

```{r}
str(data)
```

### Alternate method using sapply

```{r}
class_data <- sapply(data, class)
```
```{r}
class_data
```
- Please not that by default dataframe changes character vectors to Factors. We may turn it off using the following code:

```{r}
data <- data.frame(country = country, gdp_growth = gdp_growth, mkt_type = mkt_type, stringsAsFactors = F)
print(data)
```

```{r}
str(data)
```

